### 2017-02-22 docs/liquidcrystal.tex
Trunk add LC emulsion document and relevant changes
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@3079 40876441-3496-0410-a613-a2fca35b2ae2
Various layout and text improvements
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@3082 40876441-3496-0410-a613-a2fca35b2ae2

\section{Liquid Crystals}

@@

The bulk contributions are
\begin{equation}
f(Q_{\alpha\beta}) =
{\textstyle\frac{1}{2}} A_0 (1 - \gamma / 3) Q_{\alpha\beta}^2
-{\textstyle\frac{1}{3}} A_0 \gamma Q_{\alpha\beta} Q_{\beta\pi} Q_{\pi\alpha}
+ {\textstyle\frac{1}{4}} A_0 \gamma (Q_{\alpha\beta}^2)^2.
\end{equation}
Here, $A_0$ is a constant which sets the overall energy scale; $\gamma$
is a temperature-like parameter which controls the position in the
phase diagram relative to the isotropic-nematic transition.

@@

The gradient term in $\kappa_0$, along with the first term in $\kappa_1$
arising from the expansion of the brackets in
Equation~\ref{equation-lc-gradient-fe}, give
\begin{equation}
\kappa_0 \partial_\alpha \partial_\gamma Q_{\gamma\beta}.
+
\kappa_1 \partial_\gamma
( \partial_\gamma Q_{\alpha\beta} - \partial_\alpha Q_{\gamma\beta}).
\nonumber
\end{equation}
Note that in the one constant approximation, this term collapses
to $\kappa \partial^2 Q_{\alpha\beta}$. The cross term in the expansion
of the brackets in~\ref{equation-lc-gradient-fe}, having been symmetrised,
contributes
\begin{equation}
-2\kappa_1 q_0
( \epsilon_{\alpha\mu\nu} \partial_\mu Q_{\nu\beta}
+ \epsilon_{\beta\mu\nu} \partial_\mu Q_{\nu\alpha}),
\nonumber
\end{equation}
which must again be forced to be traceless to give
\begin{equation}
-2\kappa_1 q_0 \big\{
( \epsilon_{\alpha\mu\nu} \partial_\mu Q_{\nu\beta}
+ \epsilon_{\beta\mu\nu} \partial_\mu Q_{\nu\alpha})
- {\textstyle\frac{1}{3}}( 
   \epsilon_{\pi\mu\nu} \partial_\mu Q_{\nu\pi}
  +\epsilon_{\pi\mu\nu} \partial_\mu Q_{\nu\pi} )
   \delta_{\alpha\beta} \big\}.
\nonumber
\end{equation}
The final term in is
$-4 \kappa_1 q_0^2 Q_{\alpha\beta}$. The complete expression for the molecular
field is then
\begin{eqnarray}
H_{\alpha\beta} &=&
- A_0 (1 - \gamma/3) Q_{\alpha\beta}
+ A_0 \gamma (Q_{\alpha\pi} Q_{\pi\beta}
               - {\textstyle\frac{1}{3}} Q_{\pi\sigma}^2\delta_{\alpha\beta})
- A_0 \gamma Q_{\pi\sigma}^2 Q_{\alpha\beta}
\nonumber\\
&&
+\kappa_0 \partial_\alpha \partial_\gamma Q_{\gamma\beta}
+
\kappa_1 \partial_\gamma
( \partial_\gamma Q_{\alpha\beta} - \partial_\alpha Q_{\gamma\beta})
\nonumber\\
&&
-2\kappa_1 q_0 \big(
( \epsilon_{\alpha\mu\nu} \partial_\mu Q_{\nu\beta}
+ \epsilon_{\beta\mu\nu} \partial_\mu Q_{\nu\alpha})
- {\textstyle\frac{1}{3}}( 
   \epsilon_{\pi\mu\nu} \partial_\mu Q_{\nu\pi}
  +\epsilon_{\pi\mu\nu} \partial_\mu Q_{\nu\pi} )
   \delta_{\alpha\beta} \big)
\nonumber\\
&& - 4 \kappa_1 q_0^2 Q_{\alpha\beta}.
\end{eqnarray}

@@

\subsubsection{Normal or homeotropic anchoring}
The preferred direction of the surface order here is, as the name
suggests, normal to surface. We can write
\begin{equation}
f_s = {\textstyle\frac{1}{2}} w (Q_{\alpha\beta} - Q_{\alpha\beta}^0)^2
\end{equation}
where $w$ is a constant.
The preferred orientation $Q^0_{\alpha\beta}$ is based on the unit normal
at the surface $n_\gamma$, and is computed via the uniaxial approximation:
\begin{equation}
Q^0_{\alpha\beta}
= {\textstyle \frac{1}{2}} A (3n_\alpha n_\beta - \delta_{\alpha\beta}).
\end{equation}
The amplitude $A$ is provided by Equation~\ref{equation-lc-amplitude}. The
full boundary condition for the gradient of the tensor order parameter
at the solid-fluid boundary from Equation~\ref{equation-lc-general-bc}
is then:
\begin{eqnarray}
{\textstyle\frac{1}{2}} \kappa_0 (n_\alpha \partial_\gamma Q_{\beta\gamma}
+ n_\beta \partial_\gamma Q_{\alpha\gamma})
+ \kappa_1 n_\gamma \partial_\gamma Q_{\alpha\beta}
- {\textstyle\frac{1}{2}} \kappa_1 n_\gamma ( \partial_\alpha Q_{\gamma\beta}
+ \partial_\beta Q_{\gamma\alpha})
\nonumber
\\
- \kappa_1 q_0 n_\gamma (\epsilon_{\alpha\gamma\sigma} Q_{\sigma\beta}
+ \epsilon_{\beta\gamma\sigma}Q_{\sigma\alpha})
- w(Q_{\alpha\beta} - Q_{\alpha\beta}^0) = 0.
\label{equation-lc-bc-normal}
\end{eqnarray}

@@

\begin{figure}[h]
\begin{center}
\input{hybrid2.epic}
\end{center}
\caption{The hybrid picture. In the lattice
Boltzmann picture (left) a surface is defined by a set of links
$f_b$, which involve discrete vectors $\mathbf{c}_b \Delta t$ which
connect fluid and solid sites. For the order parameter (right), the
colloid is represented by the set of faces, e.g., that between sites
$i,j$ and $i+1,j$ with unit normal $\hat{n}_x$. This is a rather small
colloid.}
\label{figure-lc-hybrid}
\end{figure}

>>>>

\section{Liquid Crystals}
\label{section-lc}

@@

The bulk contributions are
\begin{equation}
\label{eq-lc-fed-bulk}
f(Q_{\alpha\beta}) =
{\textstyle\frac{1}{2}} A_0 (1 - \gamma / 3) Q_{\alpha\beta}^2
-{\textstyle\frac{1}{3}} A_0 \gamma Q_{\alpha\beta} Q_{\beta\pi} Q_{\pi\alpha}
+ {\textstyle\frac{1}{4}} A_0 \gamma (Q_{\alpha\beta}^2)^2.
\end{equation}
Here, $A_0$ is a constant which sets the overall energy scale; $\gamma$
is a temperature-like parameter which controls the position in the
phase diagram relative to the isotropic-nematic transition.

@@

The gradient term in $\kappa_0$, along with the first term in $\kappa_1$
arising from the expansion of the brackets in
Equation~\ref{equation-lc-gradient-fe}, give
\begin{equation}
\kappa_0 \partial_\alpha \partial_\gamma Q_{\gamma\beta}.
+
\kappa_1 \partial_\gamma
( \partial_\gamma Q_{\alpha\beta} - \partial_\alpha Q_{\gamma\beta}).
\nonumber
\end{equation}
Note that in the one constant approximation, this term collapses
to $\kappa \partial^2 Q_{\alpha\beta}$. The cross term in the expansion
of the brackets in~\ref{equation-lc-gradient-fe}, having been symmetrised,
contributes
\begin{equation}
-2\kappa_1 q_0
( \epsilon_{\alpha\mu\nu} \partial_\mu Q_{\nu\beta}
+ \epsilon_{\beta\mu\nu} \partial_\mu Q_{\nu\alpha}),
\nonumber
\end{equation}
which must again be forced to be traceless to give
\begin{equation}
-2\kappa_1 q_0 \big\{
( \epsilon_{\alpha\mu\nu} \partial_\mu Q_{\nu\beta}
+ \epsilon_{\beta\mu\nu} \partial_\mu Q_{\nu\alpha})
- {\textstyle\frac{1}{3}}( 
   \epsilon_{\pi\mu\nu} \partial_\mu Q_{\nu\pi}
  +\epsilon_{\pi\mu\nu} \partial_\mu Q_{\nu\pi} )
   \delta_{\alpha\beta} \big\}.
\nonumber
\end{equation}
The final term in is
$-4 \kappa_1 q_0^2 Q_{\alpha\beta}$. The complete expression for the molecular
field is then
\begin{eqnarray}
\label{eq-lc-h-full}
H_{\alpha\beta} &=&
- A_0 (1 - \gamma/3) Q_{\alpha\beta}
+ A_0 \gamma (Q_{\alpha\pi} Q_{\pi\beta}
               - {\textstyle\frac{1}{3}} Q_{\pi\sigma}^2\delta_{\alpha\beta})
- A_0 \gamma Q_{\pi\sigma}^2 Q_{\alpha\beta}
\nonumber\\
&&
+\kappa_0 \partial_\alpha \partial_\gamma Q_{\gamma\beta}
+
\kappa_1 \partial_\gamma
( \partial_\gamma Q_{\alpha\beta} - \partial_\alpha Q_{\gamma\beta})
\nonumber\\
&&
-2\kappa_1 q_0 \big(
( \epsilon_{\alpha\mu\nu} \partial_\mu Q_{\nu\beta}
+ \epsilon_{\beta\mu\nu} \partial_\mu Q_{\nu\alpha})
- {\textstyle\frac{1}{3}}( 
   \epsilon_{\pi\mu\nu} \partial_\mu Q_{\nu\pi}
  +\epsilon_{\pi\mu\nu} \partial_\mu Q_{\nu\pi} )
   \delta_{\alpha\beta} \big)
\nonumber\\
&& - 4 \kappa_1 q_0^2 Q_{\alpha\beta}.
\end{eqnarray}

@@

\subsubsection{Normal or homeotropic anchoring}
The preferred direction of the surface order here is, as the name
suggests, normal to surface. We can write
\begin{equation}
f_s = {\textstyle\frac{1}{2}} w (Q_{\alpha\beta} - Q_{\alpha\beta}^0)^2
\end{equation}
where $w$ is a constant.
The preferred orientation $Q^0_{\alpha\beta}$ is based on the unit normal
at the surface $n_\gamma$, and is computed via the uniaxial approximation:
\begin{equation}
Q^0_{\alpha\beta}
= {\textstyle \frac{1}{2}} A (3n_\alpha n_\beta - \delta_{\alpha\beta}).
\end{equation}
The amplitude $A$ is provided by 
\begin{equation}
\label{equation-lc-amplitude}
A=\frac{2}{3}*(\frac{1}{4}+\frac{3}{4}*\sqrt{(1 - 8/(3\gamma)})).
\end{equation}
The full boundary condition for the gradient of the tensor order parameter
at the solid-fluid boundary from Equation~\ref{equation-lc-general-bc}
is then:
\begin{eqnarray}
{\textstyle\frac{1}{2}} \kappa_0 (n_\alpha \partial_\gamma Q_{\beta\gamma}
+ n_\beta \partial_\gamma Q_{\alpha\gamma})
+ \kappa_1 n_\gamma \partial_\gamma Q_{\alpha\beta}
- {\textstyle\frac{1}{2}} \kappa_1 n_\gamma ( \partial_\alpha Q_{\gamma\beta}
+ \partial_\beta Q_{\gamma\alpha})
\nonumber
\\
- \kappa_1 q_0 n_\gamma (\epsilon_{\alpha\gamma\sigma} Q_{\sigma\beta}
+ \epsilon_{\beta\gamma\sigma}Q_{\sigma\alpha})
- w(Q_{\alpha\beta} - Q_{\alpha\beta}^0) = 0.
\label{equation-lc-bc-normal}
\end{eqnarray}

@@

\begin{figure}[h]
\begin{center}
\input{figures/hybrid.tex}
\end{center}
\caption{The hybrid picture. In the lattice
Boltzmann picture (left) a surface is defined by a set of links
$f_b$, which involve discrete vectors $\mathbf{c}_b \Delta t$ which
connect fluid and solid sites. For the order parameter (right), the
colloid is represented by the set of faces, e.g., that between sites
$i,j$ and $i+1,j$ with unit normal $\hat{n}_x$. This is a rather small
colloid.}
\label{figure-lc-hybrid}
\end{figure}
